EDIpack2: interoperable Lanczos-based exact diagonalization solver for generic quantum impurity problems.

### 2025-03-20 edipack2.tex
Updated draft. Sketched Section 3: Installation

\usepackage{listings}
\lstset{
  frame=lines,
  language=fortran,
  basicstyle=\footnotesize,
  stringstyle=\ttfamily,
  commentstyle=\itshape,
  fontadjust=true,
  keywordstyle=\color{magenta},
  mathescape,
  numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=3pt}

@@

\usepackage{bbold}
\def\11{\mathbb{1}}
\def\00{\mathbf{0}}
\def\NAME{{\rm EDIpack2.0}}

@@

\title{\NAME plus Triqs interface}
\author[]{L.~Crippa}
\author[]{I.~Krivenko}
\author[]{S.~Giuli}
\author[]{G.~Bellomia}
\author[]{....}
% \author[]{A.~Scazzola}
% \author[]{F.~Petocchi}
% \author[]{G.~Mazza}
% \author[]{L.~de Medici}
% \author[]{M.~Capone}
\author[a]{A.~Amaricci\corref{author}}

\cortext[author] {Corresponding author.\\\textit{E-mail address:} amaricci@iom.cnr.it}

\address[a]{CNR-IOM, Istituto Officina dei Materiali,
    Consiglio Nazionale delle Ricerche, Via Bonomea 265, 34136 Trieste, Italy}
\begin{abstract}

@@

% Computer program descriptions should contain the following
% PROGRAM SUMMARY.
\noindent
{\bf PROGRAM SUMMARY}
\begin{small}
  \noindent
  \\
  {\em Program Title:}  \NAME                                        \\
{\em Licensing provisions:} GPLv3\\
{\em Programming language:}  Fortran, Python \\
{\em Classification:} 6.5, 7.4, 20 \\
{\em Required dependencies:} CMake ($>=3.0.0$), Scifortran, MPI\\
{\em Nature of problem:}. \\
{\em Solution method:} .\\
\end{small}

\section{Introduction}\label{SecIntro}

\section{Library Structure}\label{SecStruct}
\subsection{Dependencies}
\subsection{EDIpack2}
\subsection{EDIpy}
\subsection{EDIpack2Triqs}

\section{Installation}\label{SecInstall}
\subsection{Building}
\subsection{Install}
\subsection{Python API and TRIQS interface}
\subsection{Conda Packaging}

\section{Usage}\label{SecUsage}
\subsection{Basic Fortran: Bethe lattice DMFT}
\subsection{Python code: Attractive Hubbard model}
\subsection{TRIQS interface: Multi-orbital Hubbard}
\subsection{Advance: Non-SU2}

\section{EDIpack2 Implementation}\label{SecEDIpack}

\subsection{The quantum impurity problem}\label{sSecQIM}
We consider a system of $N_s$ electronic levels.  A portion $N_\alpha$  of them,
i.e. the impurity levels, interact via a local repulsion, while the
remaining $N_s-N_\alpha$, i.e. the  \textit{bath} levels, are  non-interacting.
In a typical setup the impurity levels are independently coupled to a
set of $N_b$ electronic levels, so that the total number is $N_{s}=N_\alpha 
(N_{b}+1)$. Other choices for the bath topology are possible, which entail a different counting of the total levels.
The Hamiltonian of the electronic system we consider has the form:
\begin{equation}\label{Ham1}
  \begin{split}
    \hat{H}^{e}=&\hat{H}^0+\hat{H}^{int} \\
    \hat{H}^0 = &
    \sum_{\a\b\sigma}
    H^{loc}_{\a\b\sigma}d^{+}_{\a\sigma}d_{\b\sigma}  + \cr
    &\sum_{\nu\a\b\sigma}h^\nu_{\a\b\sigma}a^{+}_{\nu\a\sigma}a_{\nu\b\sigma}+ 
  % 
    \sum_{\nu\a\sigma}V^\nu_{\a\sigma}d^{+}_{\a\sigma}a_{\nu\a\sigma}+
    H.c. \\
    \hat{H}^{int}=& U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}-\\
    &J_X\sum_{\a\neq \b}d^{+}_{\a\uparrow}d_{\a\downarrow}d^{+}_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a \neq \b}d^{+}_{\a\uparrow}d^{+}_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow}
\end{split}
\end{equation}
where $a_{\alpha\sigma}$, $d_{\a\sigma}$ ($a^+_{\alpha\sigma}$,
$d^+_{\a\sigma}$) are, respectively, the destruction (creation) operators for
the bath and impurity electrons with
orbital $\alpha$ and spin $\sigma$,
$n_{\alpha\sigma}=d^{+}_{\alpha\sigma}d_{\alpha\sigma}$. 
The term $H_{\a\b\s}^{loc}$ is the non-interacting part of
the impurity Hamiltonian, 
$h_{\a\b\s}^\nu$ and $V_{\a\s}^\nu$ are, respectively, the local Hamiltonian
and the impurity hybridization of the $\nu-$th bath level. 
  Finally, $\hat{H}^{int}$ is the local
  multi-orbital interaction~\cite{Georges2013ACMP}.
  The first three terms represent the density-density part of the
  interaction, while the remaining two are,
  respectively, the spin-exchange and pair-hopping terms. 
  $U$ is the local Coulomb interaction strength, $J$ is the Hund's
  coupling~\cite{Georges2013ACMP}. We introduced
  independent coupling controlling the spin-exchange and pair-hopping terms,
  respectively $J_X$ and $J_P$. The fully symmetric Kanamori
  interaction is obtained setting $U'=U-2J$ and
  $J_X=J_P=J$~\cite{Georges2013ACMP}.

\subsection{The  basis states}\label{sSecBasis}
A natural representation of the basis states for the Fock space $\FF_e$
is obtained in the occupation number formalism of second quantization.
The \textit{Fock basis} for a finite system of $N_s$ electrons is
composed of states of the form
$\ket{\vec{n}}=\ket{n_{1\up},\dots,n_{N_s\up},n_{1\dw},\dots,n_{N_s\dw}}$, 
where each element $n_{a\sigma}=0,1$ describes the absence or the
presence of an electron with spin $\sigma$ at the level $a$. In
conjunction with the basis states, we introduce the non-Hermitian, anti-commuting, annihilation and
creation operators $c_{a\sigma}$ and
$c^{+}_{a\sigma}$, respectively. These operators act on the states
$\ket{\vec{n}}$ as:  
\begin{equation}
  \begin{aligned}
    % 1.
    c_{a\sigma}\ket{\vec{n}}=
    \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!-\!1,\dots}
      &\text{if $n_{a\sigma}\!=\! 1$}\\
      0 &\text{otherwise}
    \end{cases};
  \end{aligned}
\end{equation}
    \begin{equation}
  \begin{aligned}
    c^{+}_{a\sigma}\ket{\vec{n}}=
     \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!+\!1,\dots}
      & \text{if $n_{a\sigma}\!=\! 0$}\\
      0 & \text{otherwise}
    \end{cases}    
  \end{aligned}
\end{equation}
with $\#_{a\sigma}=\sum_{b\sigma'<a\sigma} n_{b\sigma'}$. 
Thus, each state is represented as a string of zeros and
ones, i.e. the binary decomposition of a given integer number $I$. Using
the identification $\ket{\vec{n}}=\ket{I}$, with
$I=0,\dots,2^{2N_s}-1$, each state in the Fock space can be encoded in a computer using
a sequence of $2N_s$ bits or, equivalently, an integer number in a
fixed representation.
The exponentially growing size of the Fock space will
eventually make such representation unpractical.
A solution is obtained by decomposing each state according to the
existing QNs. For a given set $[\vec{N}_\up,\vec{N}_\dw]$ of QNs we
then have: 
\begin{equation}
  \begin{split}
  \ket{\vec{n}} &=
  \prod_{\a=1}^{N_{ud}}\prod_{\sigma=\up\dw}\ket{n_1\dots
    n_{N_{bit}}}_{\a\sigma}\cr
  &=
  \begin{cases}
    \ket{\vec{n}_\up}\ket{\vec{n}_\dw},\
    \text{if}\ N_{ud}=1\\
    %
      \ket{\vec{n}_{1\up}}\cdots\ket{\vec{n}_{N_{ud}\up}}
      \ket{\vec{n}_{1\dw}}\cdots\ket{\vec{n}_{N_{ud}\dw}},\
      \text{if}\ N_{ud}=N_\alpha
    \end{cases}
    \end{split}
\end{equation}
so that, if the total number of electrons with spin $\up$ and $\dw$ is 
conserved, any state is identified by two binary sequences of
$N_{bit}=N_s$ bits, one per spin orientation.
Alternatively, if the number of electrons with spin $\up$ and $\dw$ per orbital is conserved, the
states are decomposed into two sets of
binary sequences (one set per spin orientation), each sequence
being made of  $N_{bit}=N_s/N_\alpha$ bits.
Each binary sequence is associated to a suitable tuple
of integer numbers,  univocally identifying the Fock state:
$I \rightarrow [\vec{I}_\up,\vec{I}_\dw]$, 
where $I=0,\dots,2^{2N_s}-1$ and
$\vec{I}_\sigma=[I_{1\sigma},\dots,I_{N_{ud}\sigma}]$ with
$I_{\a\sigma}=0,\dots,2^{N_{bit}}-1$. 
Through such decomposition, each state can be described by the
smallest bit set compatible with the conserved QNs.
This setup generalizes the method introduced by Lin and Gubernatis in
Ref.\onlinecite{Lin1993CIP}. 
The relation
between the Fock state index $I$ and its tuple decomposition can be
easily inverted:
\begin{equation}\label{Irelation}
I = I_1 + \sum_{i=2}^{2N_{ud}}I_i 2^{N_{bit}(i-1)}\,,
\end{equation}
where we rearranged the tuple as $[\vec{I}_\up,\vec{I}_\dw]=[I_1,\dots,I_{2N_{ud}}]$.

Such organization of the Fock states is used to construct
a suitable basis for the sectors $\SS[\vec{N}_\up,\vec{N}_\dw]$.
To any given Fock state $\ket{\vec{n}}$ and its
integer representation $I$, containing the correct bit decomposition
dictated by the QNs, it is associated a state $\ket{i}$ and an integer
$i=1,\dots,D_\SS$ through a suitable \textit{map} $\vec{\MM}_\SS$.
In particular,
each tuple of integers identifying a Fock state belonging to $\SS$ is associated to a new
tuple specific for each sector state:
$I\in\SS=[\vec{I}_\up,\vec{I}_\dw]\in\SS \xrightarrow{\vec{\MM}_\SS}
[\vec{i}_\up,\vec{i}_\dw]=i$, 
where $\vec{i}_\sigma=[i_{1\sigma},\dots,i_{N_{ud}\sigma}]$ and
$i_{\a\sigma}=1,\dots,D_{\a\sigma}$.
The tuple $[\vec{i}_\up,\vec{i}_\dw]$ univocally identifies a basis state
$\ket{i}=\ket{\vec{i}_\up,\vec{i}_\dw}$ of the sector $\SS$, through a
relation similar to \eqref{Irelation}:
\begin{equation}
i = i_{1\up} +
\sum_\sigma\sum_{l=2}^{N_{ud}}(i_{l\sigma}-1)\prod_{\a=1}^{l-1}D_{\a\sigma}
\label{SectorIndxDecomp}
\end{equation}

\subsection{Conserved quantum numbers}\label{sSecQNs}
A system of $N_s$ electrons is associated to
a Fock space of the form $\FF_e=\bigoplus_{n=0}^{N_s}
S_-\HH_e^{\otimes n}$,  where $\HH_e=\{\ket{0},\ket{\up},\ket{\dw},\ket{\up\dw} \}$ is the local
Hilbert space of the electrons on a single level,
$S_-$ is the anti-symmetrization operator, $\bigoplus$ is the direct sum and
$\bigotimes$ the tensor product.
The total dimension of the Fock space $\FF_e$ is
$\text{dim}(\HH_e)^{N_s}=4^{N_s}= 2^{2N_s}$.

The solution of the eigenvalue problem for $H_e$ is simplified by taking
into account the existence of conserved quantities.
The Hamiltonian \eqref{Ham1} conserves the total charge
$\hat{N}$ and the spin component $\hat{S}_z$, as long as we assume
that no terms  breaking the particle number or spin conservation are present,
e.g. local spin-orbital coupling, in-plane magnetic field,
superconducting order, etc.
The conservation of  $\hat{N}$ and $\hat{S}_z$ can be conveniently
re-expressed in terms of conserved number of electrons with spin $\up$ and $\dw$, i.e. $\hat{N}_\up$,
$\hat{N}_\dw$.
Moreover,  if the terms $H^{loc}$ and $h$ are
diagonal and we consider only the density-density part of the
interaction, i.e. $J_X=J_P=0$, the number of electrons with spin $\up$ and $\dw$ is conserved \textit{per
  orbital}, i.e. $\hat{N}^m_\up$, $\hat{N}^m_\dw$, where
$m=1,\dots,N_\alpha$.

%
% 
In order to formally unify the treatment of these two cases we introduce some convenient
parameters,
  namely  $N_{ud}=1$ $(N_\alpha)$ and
$N_{bit}=N_s$ $(N_s/N_\alpha)$ for conserved total (orbital resolved) number of
electrons with spin $\up$ and $\dw$.
The first corresponds to the number of conserved operators per
spin, the latter indicates the number of electronic levels per spin
(spin and orbital). In the following section we will give a more precise meaning to
$N_{bit}$.

In the rest of this work we will indicate the set of conserved quantum
numbers (QN)s with the tuple $[\vec{N}_\up,\vec{N}_\dw]$, where ($\sigma=\up,\dw$):
\begin{equation}
  \vec{N}_\sigma=
  \begin{cases}
    N_{1\sigma}\==N_\sigma, & \text{if}\ N_{ud}=1 \\
    [N_{1\sigma},\dots,N_{N_\alpha\sigma} ], & \text{if}\ N_{ud}=N_\alpha    
  \end{cases}
\end{equation}
  Note that for $N_{ud}=1$ we identified the single
  component $N_{1\sigma}$ of the vector $\vec{N}_\sigma$ with the
  total number of electrons with spin $\sigma$, i.e. $N_{\sigma}$.

In presence of a set of conserved QNs,  the Fock space decomposes into
a number of sub-spaces of reduced dimension, each
corresponding to a given value of the QNs tuple. We indicate
each sub-space with the term \textit{sector} and the symbol
$\SS[\vec{N}_\up,\vec{N}_\dw]$, or just $\SS$ where no confusion
arises.
The dimension of each sector is given by the number of ways we can
arrange $N_{\a\sigma}$ elements into $N_{bit}$ positions, i.e.:
\begin{equation}
  \begin{split}
  \text{dim}\left(\SS[\vec{N}_\up,\vec{N}_\dw]\right)  & =  \prod_{\a=1}^{N_{ud}}
  \binom{N_{bit}}{N_{\a\up}}
  \prod_{\a=1}^{N_{ud}}
  \binom{N_{bit}}{N_{\a\dw}}\cr
  &\==
  \prod_{\a=1}^{N_{ud}}D_{\a\up}
  \prod_{\a=1}^{N_{ud}}D_{\a\dw}
  \stackrel{\text{def}}{=}D_{\up}D_{\dw} \== D_\SS
  \end{split}
 \label{nud_dimension}
\end{equation}
where we introduced the symbols $D_\sigma$ and $D_{\a\sigma}$  to
indicate, respectively, the dimension of total and orbital spin-subspace 
associated to the given set of QNs. Note that $D_\sigma\==D_{1,\sigma}$ for
$N_{ud}=1$.

\subsection{Krylov based Diagonalization}\label{sSecHam}
The matrix representing the system Hamiltonian in the Fock space has a
block-diagonal structure in presence of a given set of conserved QNs.
Each block corresponds to the Hamiltonian of a sector
$\SS[\vec{N}_\up,\vec{N}_\dw]$. The analysis of the spectrum, thus,
reduces to the recursive analysis of the sector Hamiltonians. 
In each sector, the Hamiltonian of the system has the following general form: 
\begin{equation}
H^e_\SS = H_d  + H_\up\otimes \mathbb{I}_\dw + \mathbb{I}_\up\otimes
H_\dw + H_{nd}\,.
\label{Hdecomp}
\end{equation}
$H_d$ is a diagonal term containing the local part of the Hamiltonian,
including the density-density terms of the interaction.
The $H_\sigma$ components describe all the hopping processes of the
electrons with spin $\sigma=\up,\dw$.
Finally the term $H_{nd}$ contains all the remaining non-diagonal
elements which do not fit in the previous two components,
e.g. the spin-exchange and pair-hopping interaction terms.
If the QNs are conserved per orbital, i.e. if $H_{nd}\==0$ and no
inter-orbital local hopping terms are present,  each $H_\sigma$
further splits into a sum of smaller terms:
$$
H_\sigma = \sum_{\a=1}^{N_\alpha}\mathbb{I}_{1\sigma}\otimes\cdots\otimes
H_{\a\sigma}\otimes\cdots\otimes \mathbb{I}_{N_\alpha\sigma}
$$

Each term of the  Hamiltonian matrix is constructed  independently 
iterating over one or more  components of the sector basis states.
For example, the construction of the matrices $H_{\a\sigma}$ only involve iterations over the components
$\ket{i_{\a\sigma}}$ of the sector basis. In general we have:
\begin{equation}
\begin{split}
  \bra{\vec{i}_\up\vec{i}_\dw}H^e_\SS\ket{\vec{j}_\up\vec{j}_\dw}
  =&
  \bra{\vec{i}_\up\vec{i}_\dw}
  H_d + H_\up\otimes \mathbb{I}_\dw + \mathbb{I}_\up\otimes H_\dw + H_{nd}
  \ket{\vec{j}_\up\vec{j}_\dw}\cr
  =&
  \bra{\vec{i}_\up\vec{i}_\dw}H_d\ket{\vec{i}_\up\vec{i}_\dw} +\cr
  &\bra{\vec{i}_\up}H_\up\ket{\vec{j}_\up}\delta_{\vec{i}_\dw\vec{j}_\dw}
  +
  \bra{\vec{i}_\dw}H_\dw\ket{\vec{j}_\dw}\delta_{\vec{i}_\up\vec{j}_\up} + \cr
  &\bra{\vec{i}_\up\vec{i}_\dw}H_{nd}\ket{\vec{j}_\up\vec{j}_\dw} 
\end{split}
\label{Hbuild}
\end{equation}

For very large systems, storing the Hamiltonian matrix in the memory
can be highly inefficient.
In such cases, Krylov or Lanczos
methods~\cite{Lanczos1950JRNBSB,Lin1993CIP,Lehoucq1998,Maschhoff1996} can
be implemented using a storage-free algorithm, performing the
necessary linear operations on-the-fly.
This solution has generally a negative impact on the execution
time, however this can be well compensated by scaling in a massively
parallel framework.

\subsection{Dynamical correlation functions, observables and reduced
  impurity density matrix}

@@

\subsection{Bath $\chi^2$ fit}

@@

\section{EDIpy: Python API}\label{SecEDIpy}

\section{edipack2triqs: a thin interface to TRIQS}\label{Secedi2triqs}

@@

% \section*{Appendix A: Input Variables}\label{AppendixA}
% Here is reported a list of the global input variables of the
% library. The input parameters are arranged into the following table,
% as discussed in \secu{SecWorkflow}.
% Note that only a subset of the input parameters are transparent to the
% user, i.e. can be directly accessed from the calling program.
% These variables  are indicated with a symbol in the first column of
% the table.  In the columns from $2^{\rm nd}$ to $4^{\rm th}$ we indicate,
% respectively, the variable name, the type and the default value.
% Finally, In the last column we provide a short
% description of the variable purpose.

% % \begin{longtable}{ M{0.01pt} T{0.19\textwidth}  T{0.04\textwidth}
%     %     D{0.11\textwidth} M{0.52\textwidth}  }

% \begin{tabularx}{\linewidth}{ M{0.01\textwidth} T{0.19\textwidth}
%     T{0.04\textwidth} D{0.11\textwidth} M{0.52\textwidth} }
%   \caption{List of the \NAME\ input variables.}\\
%   \hline 
%   \hline
%   {\it \footnotesize \phantom{T}} & {\it Variable} & {\it Type} & {\it Default} & {\it
%                                                          \footnotesize
%                                                          Description} \\
%   \hline
%   \hline
%   \endhead
%   \hline
%   \multicolumn{5}{c}{\bf Most relevant control variables}\\
%   \hline
%   $\circ$ & Norb & Int & 1 & Number of impurity orbitals.\\
%   \hline
%   $\circ$ & Nspin& Int & 1 & Number of spin channels treated
%                              independently. \texttt{Nspin}=1 enforces
%                              non-magnetic solution.\\
%   \hline
%    & Nbath& Int & 6  & Number of bath sites. See \texttt{bath\_type}
%                        for information about resulting total number of
%                        electronic levels $N_s$.\\
%   \hline
%   $\circ$ & bath\_type& Char & ``normal''    & Set the bath type:
%                                      \texttt{normal}:
%                                      $N_s=\mathtt{Norb}(1+\mathtt{Nbath})$,
%                                      \texttt{hybrid}:
%                                      $N_s=\mathtt{Norb}+\mathtt{Nbath}$,
%                                      and \texttt{replica}:
%                                      $N_s=\mathtt{Norb}(1+\mathtt{Nbath})$,
%                                      See \secu{SecBath}.\\
%   \hline
% %--------
%   $\circ$ & Uloc& Dble & 2.0   & Local intra-orbital interactions $U$,
%                                      different values per orbitals can
%                                      be defined.\\
%   \hline
%   $\circ$ & Ust& Dble & 0.0   & Local inter-orbital interaction $U'$,
%                                 with opposite spin orientations.
%                                 For the Kanamori interaction this 
%                                 must be set to
%                                 $\mathtt{Uloc}-2\mathtt{Jh}$, with \texttt{Jh}
%                                 the Hund's coupling. 
%                                 The value $U''$ for equal spin orientation
%                                 is automatically set by the difference 
%                                 $\mathtt{Ust}-\mathtt{Jh}$. \\
%   \hline
%   $\circ$ & Jh   & Dble & 0.0 &  Hund's coupling.\\
%   \hline
%   $\circ$ & Jx   & Dble & 0.0 & Spin-Exchange coupling constant.\\
%   \hline
%   $\circ$ & Jp   & Dble & 0.0 & Pair-Hopping coupling constant.\\
%   \hline
%     $\circ$ & beta   & Dble & 1000 & Inverse temperature. For $T=0$
%                                    calculations this value sets a 
%                                    discretization of the Matsubara
%                                    frequencies.\\
%   \hline
%   $\circ$ & xmu   & Dble & 0.0 & Chemical potential. This parameter
%                                  sets the average occupations of the
%                                  impurity. 
%                                  If \texttt{HFMODE=True} this value
%                                  contains the Hartree shift, i.e. $\mathtt{xmu}=0.0$ sets the half-filling for a particle-hole
%                                  symmetric case.\\
%   \hline
%      & HFmode  & Bool & .true. & If True the Hartree shift is included in the
%                 interaction, e.g. $U(n_\up-1/2)(n_\dw-1/2)$. In this
%                        case zero chemical potential corresponds to
%                        half-filling for a particle-hole symmetric system.\\  
%   \hline
%   $\circ$ & Lmats & Int & 4096 & Number of Matsubara frequencies.\\
%   \hline
%   $\circ$ & Lreal  & Int & 5000 & Number of real-axis frequencies\\
%   \hline
%     $\circ$ & eps   & Dble & 0.01 & Real-axis broadening.\\
%   \hline
%   $\circ$ & wini, wfin & Dble & -5, 5 & Real-axis frequency range.\\
%   \hline
%   $\circ$ & Nloop    & Int & 100 & Maximum number of allowed iterations in the DMFT solution.\\
%   \hline
%   $\circ$ & dmft\_error & Dble & 1e-5 & Convergence threshold for the
%                                         DMFT iterative cycle.\\
%   \hline
%   %--------
%   $\circ$ & Nph    & Int & 0 & Maximum number of phonons.\\
%   \hline
%   $\circ$ & g\_ph   & Dble & 0.0& Electron-Phonon coupling.\\
%   \hline
%   $\circ$ & w0\_ph  & Dble & 0.0& Holstein phonon frequency.\\
%   \hline
%   $\circ$ & xmin, xmax & Dble & -3, 3& Phonon probability distribution spatial range.\\
%   \hline
% %--------
%    & chispin\_flag & Bool & .false. & Flag to include the evaluation of Spin-Spin susceptibilities.\\
%   \hline
%    &  chidens\_flag &Bool & .false. & Flag to include the evaluation of Density-Density susceptibilities.\\
%   \hline
%    & chipair\_flag & Bool & .false. &  Flag to include the evaluation
%                                       of $s$-wave Pair susceptibilities.\\
%   \hline
%    & chiexct\_flag & Bool & .false. & Flag to include the evaluation of Excitonic susceptibilities.\\
%   \hline
%   %-------
%    & ed\_sparse\_H    & Bool & .true. &  Flag to select storage type of the sparse
%                          Hamiltonian. True: sparse matrix object,
%                          False:on-the-fly matrix-vector product.\\
%   \hline
%    & ed\_total\_ud     & Bool & .true. & Flag to select quantum numbers type.
%                                  True: conserved total electron occupation per
%                                  spin. 
%                                  False: conserved electrons
%                                  occupations per orbital and spin.\\
%   \hline 
%    & ed\_print\_Sigma & Bool & .true. & Flag to print out the impurity Self-energies.\\
%   \hline
%    & ed\_print\_G        & Bool & .true. & Flag to print out the interacting impurity Green's functions.\\
%   \hline
%    & ed\_print\_G0      & Bool & .true. & Flag to print out the non-interacting
%                               impurity Green's functions.\\
%   \hline
%    & ed\_twin             & Bool & .false. & Flag to reduce the number of visited
%                                        sector using  sectors symmetry:
%                                        $[\vec{N}_\up,\vec{N}_\dw]\leftrightarrow[\vec{N}_\dw,\vec{N}_\up]$.
%                                        A warning is raised if
%                                        \texttt{ed\_twin=.true.} and
%                                        \texttt{Nspin=2} as this
%                                        symmetry 
%                                        conflicts with possible long-range
%                                        magnetic order.\\ 
%   \hline
%    & ed\_verbose       & Int & 3 & Verbosity level [0-5]. \\
%   \hline
%   % ----------
%    & lanc\_method     & Char & ``arpack'' & String to select Lanczos
%                                    method. \texttt{ARPACK}: uses
%                                    P-Arpack, \texttt{LANCZOS}: uses
%                                    single vector Lanczos method
%                                    with no re-orthogonalization \\
%   \hline
%    & lanc\_tolerance  & Dble & 1e-18 & Tolerance for the Lanczos algorithm. \\
%   \hline
%    & lanc\_niter         & Int & 512 & Largest number of Lanczos
%                                        iterations used in the
%                                        Diagonalization of each sector Hamiltonian.\\
%   \hline
%    & lanc\_ngfiter      & Int & 200 & Largest dimension of the Krylov basis in
%                               the evaluation of the Green's function.\\
%   \hline

>>>>

\usepackage{listings}
\lstdefinestyle{fstyle}{
  frame=lines,
  language=fortran,
  basicstyle=\footnotesize,
  stringstyle=\ttfamily,
  commentstyle=\itshape,
  fontadjust=true,
  keywordstyle=\color{magenta},
  mathescape,
  numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=3pt
}

@@

\usepackage{bbold}
\def\11{\mathbb{1}}
\def\00{\mathbf{0}}
\def\NAME{{\rm EDIpack2 }}

@@

\title{EDIpack2: interoperable Lanczos-based exact diagonalization solver for
  generic quantum impurity problems.}
% \author[]{L.~Crippa}
% \author[]{I.~Krivenko}
% \author[]{S.~Giuli}
% \author[]{G.~Bellomia}
% \author[]{....}
% % \author[]{A.~Scazzola}
% % \author[]{F.~Petocchi}
% % \author[]{G.~Mazza}
% % \author[]{L.~de Medici}
% % \author[]{M.~Capone}
% \author[a]{A.~Amaricci\corref{author}}

% \cortext[author] {Corresponding author.\\\textit{E-mail address:} amaricci@iom.cnr.it}

% \address[a]{CNR-IOM, Istituto Officina dei Materiali,
%     Consiglio Nazionale delle Ricerche, Via Bonomea 265, 34136 Trieste, Italy}
\begin{abstract}

@@

% Computer program descriptions should contain the following
% PROGRAM SUMMARY.
\noindent
{\bf PROGRAM SUMMARY}
\begin{small}
  \noindent
  \\
  {\em Program Title:}  EDIpack2                                        \\
{\em Licensing provisions:} GPLv3\\
{\em Programming language:}  Fortran, Python \\
{\em Classification:} 6.5, 7.4, 20 \\
{\em Required dependencies:} CMake ($>=3.0.0$), Scifortran, MPI\\
{\em Nature of problem:}. \\
{\em Solution method:} .\\
\end{small}

\section{Introduction and Motivation}\label{SecIntro}

A few words about the motivations who led us to develop this
software, possible applications and advantages:

* flexibility: can address generic cases including multi-orbital,
superconducting or spin-non-conserving regimes

* zero and low finite temperatures

* direct access to (well approximated) analytic dynamical functions

* direct access to impurity Fock space quantities

bla bla bla

\section{Structure}\label{SecStruct}
\subsection{EDIpack2}
\subsection{EDIpack2ineq}
\subsection{EDIpack2 C-bindings}
\subsection{EDIpack2Triqs}

\section{Installation}\label{SecInstall}
The installation of \NAME is available through CMake which ensures
multi-platforms compabitility and dependencies resolution.  
The software builds into two distinct libraries: {\tt
  libedipack.a} and {\tt libedipack\_cbinding.so}.
The former, alongside the generated Fortan modules files,
wraps the \NAME software. The latter enables interoperability through
bindings to the C language.

\subsection{Dependencies}
Although \NAME is as much self-contained as possible its development
hinges on two external libraries.
\begin{itemize}
\item {\bf SciFortran}: an open-source Fortran library to support
  mathematical and scientific software development. 
\item {\bf MPI} (optional): a distributed memory parallel communication layer with support to modern Fortran compiler.
\end{itemize}

SciFortran provides a solid development platform enabling access to
many algorithms and functions, including standard linear algebra
operations and high-performance Lanczos based algorithms. This
greatly reduces code clutter and development time.
The use of distributed memory parallel environment, although optional,
is required to access scalable parallel diagonalization algorithms
which speed up calculations for large dimensional systems.

\subsection{Build and Install}
\subsubsection{Source}
The software can be installed from source as follows. The source can
be retrieved directly from its GitHub repository, for instance using:
\begin{lstlisting}[language=Bash]
git clone https://github.com/edipack/EDIpack2.0 EDIpack2
\end{lstlisting}
Then, assuming to be in the software directoru, a conventional
out-of-source building is performed using two different compilations
backends.

\begin{itemize}
  \item {\bf GNU Make}\\
This is the default CMake workflow:
\begin{lstlisting}[language=Bash]
mkdir build
cd build
cmake ..
make -j
make install
\end{lstlisting}

\item{\bf Ninja}

An alternative workflow employs the Ninja building backend with
Fortran support. Ninja is generally faster and automatically supports
multi-threaded building:
\begin{lstlisting}[language=Bash]
mkdir build
cd build
cmake -GNinja ..
ninja
ninja install
\end{lstlisting}
\end{itemize}

\noindent
The CMake configurations can be further tuned using the following variables:
\begin{center}
\begin{tabular}{ l|l|l } 
 \hline
  {\bf Option}               & {\bf Scope} & {\bf Value} \\
  \hline
  -D{\bf PREFIX}          & Install directory  & $\sim$/opt/EDIpack2/TAG/PLAT/BRANCH\\
  -D{\bf USE\_MPI}       & MPI support  &  True/{\color{red}False}\\
  -D{\bf WITH\_INEQ}   & Inequivalent impurities support & {\color{red}True}/{False}\\
  -D{\bf VERBOSE}      & Verbose CMake output & {\color{red}True}/{False}\\ 
  -D{\bf BUILD\_TYPE} & Compilation flags & {\color{red}RELEASE}/TESTING/DEBUG/AGGRESSIVE \\
 \hline
\end{tabular}
\end{center}

The default target builds and install either the main library and the C-binding.
However, a specific building for each library is available specifying
the required target. A recap message is printed at the end of the
CMake configuration step.

\subsubsection{Anaconda}
As an alternative we provide for both Linux and OSx systems
installation through Anaconda packages into a virtual
environment containing Python ($>3.10$).

The Conda package installation procedure reads:
\begin{lstlisting}[language=Bash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2
\end{lstlisting}
\noindent
which installs a bundle of Scifor and \NAME libraries together with
specific {\tt pkg-config} configurations files which can be used to
retrieve compilation and linking flags.

\subsection{OS Loading}
In oder to avoid possible conflicts or require administrative
privileges, the building step results get installed by default in a user home
directory, specified by the CMake variable {\tt PREFIX}.
In doing so, however, one misses the chance of automatic loading into
the operative system.

We offer different strategies to perform this action:
\begin{enumerate}
\item  A CMake generated configuration file for environment module
  which allows to load and unload the library at any time. This is
  preferred solution for HPC systems. 
\item A CMake generated bash script to be sourced (once or
  permanently) in any shell session to add \NAME library to the
  default environment.
\item A CMake generated pkg-config configuration file to be added in
  the pkg-config path itself.  
\end{enumerate}
An automatically generated recap message with all instructions is
generated at the end of the installation procedure.

\subsection{Python API}
\subsubsection{Build from source}
EDIpy2, i.e. the Python API of EDIpack2, is available as a stand-alone
module which depends on both \NAME and SciFortran. The package can be
obtained from the repository
\href{https://github.com/EDIpack/EDIpy2.0}{EDIpy2}.

\begin{lstlisting}[language=Bash]
git clone https://github.com/edipack/EDIpy2 EDIpy2
cd EDIpy2
pip install . 
\end{lstlisting}
In some more recent Python distribution the flag {\tt
  --break-system-packages} might be required to complete
installation or a virtual environment should be used instead.

\subsubsection{Anaconda}
As for EDIpack2, also the Python API in EDIpy2 are available through
Anaconda packaging. In this case the resolution of the dependencies is
taken care from Conda itself:

\begin{lstlisting}[language=Bash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2
\end{lstlisting}

\subsection{TRIQS interface}
A purely Python \NAME to Triqs interface is available, leveraging on
the C-bindings and Python API. The corresponding module depends on
\NAME (which ultimately depends on SciFortran) and Triqs.
Assuming the two software are correctly installed in the OS, the
EDIpack2Triqs interface is installed as follows:

\begin{lstlisting}[language=Bash]
git clone https://github.com/krivenko/edipack2triqs
cd edipack2triqs
pip install .
\end{lstlisting}

\section{Usage}\label{SecUsage}
\subsection{Bethe lattice DMFT (Fortran API)}
\subsection{Attractive Hubbard model (Python API)}
\subsection{Multi-orbital Hubbard (Triqs)}

\section{Implementation}\label{SecEDIpack}

\subsection{The quantum impurity problem}\label{sSecQIM}

\subsection{The  basis states}\label{sSecBasis}

\subsection{Conserved quantum numbers}\label{sSecQNs}

\subsection{Krylov based Diagonalization}\label{sSecHam}

\subsection{Dynamical correlation functions}

\subsection{Observables}

\subsection{Reduced impurity density matrix}

@@

\subsection{Bath Optimization}

@@

\section{Inequivalent impurities}

\section{C-bindings}\label{SecCbind}

\section{Python API}

\section{Triqs interface}\label{SecEDI2Triqs}

@@
